---
title: "Mini Project #1"
format: html
---

## Importing Data

```{r}
#| warning: false

library(tidyverse)
library(dplyr)
library(sf)
library(tmap)
library(maps)
library(viridis)
library(htmltools)
library(glue)
library(leaflet)

us_states <- map_data("state")

frogs <- read.csv("~/Desktop/15/SDS264/data/most_common_frog.csv") |>
  na.omit()

forests <- read.csv("~/Desktop/15/SDS264/data/forests.csv") |>
  mutate('forest_cover' = forest_area/land_area * 100)

states <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson")
```

## Data Source

# Forest

For my numerical maps, I looked at the percentage of land covered in forest in the united states. I found this data in the USDA Forest Service FIA Annual Report, this report is a pdf where I extracted the values that I would need for this data set.

# Frogs

I wanted to look frogs for my categorical data. I found observations of frogs on the website, inaturalist where you can export observations based on specific criteria. I wanted to look at something more interesting then just the most common frog in each state so i looked at the most commonly observed frog that is deemed threatened.

## Joining for Interactive Polygons

```{r}
#| warning: false
states <- states |>
  filter(!(name %in% c("Alaska", "Hawaii", "Puerto Rico"))) |>
  select("name", "geometry")

frogs_interactive <- frogs |>
  right_join(states, by = c("state" = "name"))
```

```{r}
#| warning: false
forest_interactive <- forests |>
  right_join(states, by = c("state" = "name"))
```

## Interactive Maps

```{r}
#| warning: false
forest_sf <- st_as_sf(forest_interactive) |>
  mutate(forest_cover = trunc(forest_cover))

pal <- colorNumeric("Greens", domain = forest_sf$forest_cover)

forest_sf <- forest_sf |>
  mutate(labels = str_c(state, ": ", forest_cover, "% forest cover"))

labels <- lapply(forest_sf$labels, HTML)
leaflet(forest_sf) |>
  setView(-96, 37.8, 4) |>
  addProviderTiles("Esri.WorldTopoMap") |>
  addPolygons(
    fillColor = ~pal(forest_cover),
    weight = 2,
    opacity = 1,
    color = "black",
    fillOpacity = 0.6,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "pink",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) |>
  addLegend(pal = pal, title = "% Forest Coverage", values = ~forest_cover, opacity = 0.7, position = "bottomright") |>
  addScaleBar(position = "bottomleft") |>
  addPopups(-95, 50, "Percentage of Forest Cover in Each State",
              options = popupOptions(closeOnClick = FALSE))
```

```{r}
#| warning: false
frog_sf <- st_as_sf(frogs_interactive) 
pal <- colorFactor(c("#800000", 
                               "#9A6324", 
                               "#808000", 
                               "#469990", 
                               "#000075", 
                               "#000000", 
                               "#e6194B", 
                               "#f58231", 
                               "#ffe119", 
                               "#bfef45", 
                               "#3cb44b", 
                               "#42d4f4", 
                               "#4363d8",
                               "#911eb4", 
                               "#f032e6", 
                               "#fabed4", 
                               "#ffd8b1",
                               "#fffac8",
                               "#aaffc3"), domain = frog_sf$frog)

frog_sf <- frog_sf |>
  mutate(labels = str_c("The ", frog, " is the most observed threatened frog in ", state))

labels <- lapply(frog_sf$labels, HTML)

leaflet(frog_sf) |>
  setView(-96, 37.8, 4) |>
  addProviderTiles("Esri.WorldTopoMap") |>
  addPolygons(
    fillColor = ~pal(frog),
    weight = 2,
    opacity = 1,
    color = "black",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "pink",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "12px",
      direction = "auto")) |>
  addLegend(pal = pal, title = "Frogs Observed", values = ~frog, opacity = 0.7, position = "bottomright") |>
  addScaleBar(position = "bottomleft")
```
